Family Income and Educational Returns in China, 2010-2020: Higher Returns Favor the Wealthy, but Gaps Hold Steady

### 2024-12-04 paper/paper.qmd
updated paper and references

library(tidyverse)
library(caret)
library(glmnet)
library(nnet)
library(stringr)
library(gt)
library(broom)
library(knitr)
library(MatchIt)
library(arrow)
```

@@

I used the statistical programming language R [@R] for the graphing, analysis and presentation of the project as a whole. Caret [@caret], glmnet [@glmnet; @glmnet2; @glmnet3], nnet [@nnet], tidyverse [@tidyverse], car [@car] MatchIt [@matchit] and arrow [@arrow] were used directly in the analysis of the data. Stringr [@stringr], dplyr [@dplyr], gt [@gt], knitr [@knitr], and styler [@styler] were used in the presentation and/or styling of the data, graphs, and paper as a whole.

@@

After cleaning the data to split results by candidate, removing polls with N/A values and ratings lower than 2.5, and further selecting polls polling likely voters as opposed to the whole population, we are left with 746 data points with the variables pollster, has_sponsor, numeric_grade, pollscore, transparency_score, sample_size, end_date, state, candidate_name, pct (support rate in percentage), and cycle (election cycle, i.e. 2020/2024).

@@

Table 2 provides a summary of balance for the match between the Quartile 1 and the Quartile 2 subset of CFPS 2020. The match was executed using caliper = 0.1 and a ratio of 2, matching 1041 Quartile 2 instances out of 1324 total instances with all 541 Quartile 1 instances. This matching procedure was repeated 6 times to compare each quartile with every other quartile for all 6 CFPS iterations from 2010 to 2020. The standard mean differences of matches using caliper = 0.2 and 0.5 were included as a comparison. The caliper of 0.1 returned the best standardized mean difference of 0.0139 between propensity scores of Q1 and Q2. Caliper widths of 0.2 and 0.5 returned 0.0338 and 0.0795, respectively.

TABLE HERE IF TIME

@@

@@

Finally, the Delta coefficient $\alpha_1$ from all paired comparisons across every CFPS iteration were compiled together. 5% t-tests were carried out on the data from each iteration to discern if the differences in educational returns between quartiles were statistically significant.

@@

# Regional polls analysis
trump_2024_regional <- trump_2024_lower %>% filter(state != "National")
harris_2024_regional <- harris_2024_lower %>% filter(state != "National")

trump_Pennsylvania <- trump_2024_regional %>% filter(state == "Pennsylvania")
trump_Georgia <- trump_2024_regional %>% filter(state == "Georgia")
trump_North_Carolina <- trump_2024_regional %>% filter(state == "North Carolina")
trump_Michigan <- trump_2024_regional %>% filter(state == "Michigan")
trump_Arizona <- trump_2024_regional %>% filter(state == "Arizona")
trump_Wisconsin <- trump_2024_regional %>% filter(state == "Wisconsin")
trump_Nevada <- trump_2024_regional %>% filter(state == "Nevada")

harris_Pennsylvania <- harris_2024_regional %>% filter(state == "Pennsylvania")
harris_Georgia <- harris_2024_regional %>% filter(state == "Georgia")
harris_North_Carolina <- harris_2024_regional %>% filter(state == "North Carolina")
harris_Michigan <- harris_2024_regional %>% filter(state == "Michigan")
harris_Arizona <- harris_2024_regional %>% filter(state == "Arizona")
harris_Wisconsin <- harris_2024_regional %>% filter(state == "Wisconsin")
harris_Nevada <- harris_2024_regional %>% filter(state == "Nevada")

combined_Pennsylvania <- full_join(
  trump_Pennsylvania,
  harris_Pennsylvania,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Georgia <- full_join(
  trump_Georgia,
  harris_Georgia,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_North_Carolina <- full_join(
  trump_North_Carolina,
  harris_North_Carolina,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Michigan <- full_join(
  trump_Michigan,
  harris_Michigan,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Arizona <- full_join(
  trump_Arizona,
  harris_Arizona,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Wisconsin <- full_join(
  trump_Wisconsin,
  harris_Wisconsin,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_Nevada <- full_join(
  trump_Nevada,
  harris_Nevada,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

# mutate harris_win variable by comparing polled support rates
# preparing data for logistic regression
combined_Pennsylvania <- combined_Pennsylvania %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Georgia <- combined_Georgia %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_North_Carolina <- combined_North_Carolina %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Michigan <- combined_Michigan %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Arizona <- combined_Arizona %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Wisconsin <- combined_Wisconsin %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

combined_Nevada <- combined_Nevada %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

# Create 7 different logistic regressions to predict Harris's win rate
# in each state
logistic_Pennsylvania <- readRDS(here::here("models/Logistic_model_Pennsylvania.rds"))

logistic_Georgia <- readRDS(here::here("models/Logistic_model_Georgia.rds"))

logistic_North_Carolina <- readRDS(here::here("models/Logistic_model_North_Carolina.rds"))

logistic_Michigan <- readRDS(here::here("models/Logistic_model_Michigan.rds"))

logistic_Arizona <- readRDS(here::here("models/Logistic_model_Arizona.rds"))

logistic_Wisconsin <- readRDS(here::here("models/Logistic_model_Wisconsin.rds"))

logistic_Nevada <- readRDS(here::here("models/Logistic_model_Nevada.rds"))

# test logistic models for state specific analysis
# mutate fitted values

combined_Pennsylvania <- combined_Pennsylvania %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Pennsylvania, type = "response"), 2)
  )

roc_obj_Pennsylvania <- roc(
  combined_Pennsylvania$harris_win,
  combined_Pennsylvania$harris_win_prob
)

auc_value_Pennsylvania <- auc(roc_obj_Pennsylvania)

combined_Georgia <- combined_Georgia %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Georgia, type = "response"), 2)
  )

roc_obj_Georgia <- roc(
  combined_Georgia$harris_win,
  combined_Georgia$harris_win_prob
)

auc_value_Georgia <- auc(roc_obj_Georgia)

combined_North_Carolina <- combined_North_Carolina %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_North_Carolina, type = "response"), 2)
  )

roc_obj_North_Carolina <- roc(
  combined_North_Carolina$harris_win,
  combined_North_Carolina$harris_win_prob
)

auc_value_North_Carolina <- auc(roc_obj_North_Carolina)

combined_Michigan <- combined_Michigan %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Michigan, type = "response"), 2)
  )

roc_obj_Michigan <- roc(
  combined_Michigan$harris_win,
  combined_Michigan$harris_win_prob
)

auc_value_Michigan <- auc(roc_obj_Michigan)

combined_Arizona <- combined_Arizona %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Arizona, type = "response"), 2)
  )

roc_obj_Arizona <- roc(
  combined_Arizona$harris_win,
  combined_Arizona$harris_win_prob
)

auc_value_Arizona <- auc(roc_obj_Arizona)

combined_Wisconsin <- combined_Wisconsin %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Wisconsin, type = "response"), 2)
  )

roc_obj_Wisconsin <- roc(
  combined_Wisconsin$harris_win,
  combined_Wisconsin$harris_win_prob
)

auc_value_Wisconsin <- auc(roc_obj_Wisconsin)

combined_Nevada <- combined_Nevada %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Nevada, type = "response"), 2)
  )

roc_obj_Nevada <- roc(
  combined_Nevada$harris_win,
  combined_Nevada$harris_win_prob
)

auc_value_Nevada <- auc(roc_obj_Nevada)
```

We identified the AUC values and ROC plots for each of the seven swing-state logistic regressions. The ROC plots are presented in the Additional Data Details section (@sec-model-details), from @fig-pen to @fig-nev. The AUC values are: 0.90 for Pennsylvania, 0.88 for Georgia, 0.74 for North Carolina, 0.93 for Michigan, 1.00 for Arizona, 0.86 for Wisconsin, and 0.65 for Nevada. All states other than Nevada have acceptable or excellent AUC values, indicating that the respective logistic models are reliable when distinguishing cases where Harris wins. The low AUC value for Nevada is potentially caused by the limited sample size; there are only 9 polls conducted in Nevada.

Moreover, the sample size restrictions were the reason why we did not use training and testing data to validate our logistic models. The state with the greatest number of polls, Pennsylvania, only records 26 polls. Utilizing training and testing datasets implies that the datasets may include only a couple of data points in some states, which is insufficient to construct or test the model. On top of that, according to [@trainingerror], achieving low training error, such as by using a training-MSE-minimising regression model such as our own, is sufficient to ensure the predictive capabilities of a model bounded by some constant variance term.

# find biases of pollsters from 2020 national polls data
# filter out national polls from the datasets
trump_2024_national <- trump_2024 %>% filter(state == "National")
harris_2024_national <- harris_2024 %>% filter(state == "National")
trump_2020_national <- trump_2020 %>% filter(state == "National")
biden_2020_national <- biden_2020 %>% filter(state == "National")

# Define actual results
actual_results_2020 <- list(Trump_pct = 47, Biden_pct = 51)

# Calculate bias per pollster
rep_pollster_bias <- trump_2020_national %>%
  group_by(pollster) %>%
  summarize(
    Trump_bias = mean(pct - actual_results_2020$Trump_pct, na.rm = TRUE)
  )

dem_pollster_bias <- biden_2020_national %>%
  group_by(pollster) %>%
  summarize(
    Biden_bias = mean(pct - actual_results_2020$Biden_pct, na.rm = TRUE)
  )

# adjust 2024 national polls percentages by the average pollster biases
# also mutate has_sponsor variable into dummy variable and remove NA's
trump_2024_national_adj <- trump_2024_national %>%
  left_join(rep_pollster_bias, by = "pollster") %>%
  mutate(
    pct_adj = pct - ifelse(!is.na(Trump_bias), Trump_bias, 0)
  ) %>%
  select(-Trump_bias)

# Adjust Harris 2024 National Polls
harris_2024_national_adj <- harris_2024_national %>%
  left_join(dem_pollster_bias, by = "pollster") %>%
  mutate(
    pct_adj = pct - ifelse(!is.na(Biden_bias), Biden_bias, 0)
  ) %>%
  select(-Biden_bias)

# model adjusted support percentage by time and control variables
# first split training and testing dataset using an 80/20 split
set.seed(11451)
trainIndex_trump_adj <- createDataPartition(trump_2024_national_adj$pct_adj,
  p = 0.8, list = FALSE
)
data_train_trump_adj <- trump_2024_national_adj[trainIndex_trump_adj, ]
data_test_trump_adj <- trump_2024_national_adj[-trainIndex_trump_adj, ]

trainIndex_harris_adj <- createDataPartition(harris_2024_national_adj$pct_adj,
  p = 0.8, list = FALSE
)
data_train_harris_adj <- harris_2024_national_adj[trainIndex_harris_adj, ]
data_test_harris_adj <- harris_2024_national_adj[-trainIndex_harris_adj, ]

model_date_harris_adj <- lm(pct_adj ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_harris_adj)
model_date_trump_adj <- lm(pct_adj ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_trump_adj)

# repeat for unadjusted support levels
trainIndex_trump <- createDataPartition(trump_2024_national$pct,
  p = 0.8, list = FALSE
)
data_train_trump <- trump_2024_national[trainIndex_trump, ]
data_test_trump <- trump_2024_national[-trainIndex_trump, ]

trainIndex_harris <- createDataPartition(harris_2024_national$pct,
  p = 0.8, list = FALSE
)
data_train_harris <- harris_2024_national[trainIndex_harris, ]
data_test_harris <- harris_2024_national[-trainIndex_harris, ]

model_date_harris <- lm(pct ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_harris)
model_date_trump <- lm(pct ~ end_date + has_sponsor +
  transparency_score + sample_size, data = data_train_trump)

# Functions to evaluate model using testing dataset
evaluate_model_adj <- function(model, test_data) {
  predictions <- predict(model, newdata = test_data)
  actuals <- test_data$pct_adj
  rmse <- sqrt(mean((predictions - actuals)^2, na.rm = TRUE))
  r_squared <- summary(model)$r.squared
  return(list(RMSE = rmse, R_squared = r_squared))
}

evaluate_model <- function(model, test_data) {
  predictions <- predict(model, newdata = test_data)
  actuals <- test_data$pct
  rmse <- sqrt(mean((predictions - actuals)^2, na.rm = TRUE))
  r_squared <- summary(model)$r.squared
  return(list(RMSE = rmse, R_squared = r_squared))
}

# Evaluate Harris model with adjusted support percentages
harris_evaluation_adj <- evaluate_model_adj(model_date_harris_adj, data_test_harris_adj)

# Evaluate Trump model with adjusted support percentages
trump_evaluation_adj <- evaluate_model_adj(model_date_trump_adj, data_test_trump_adj)

# Evaluate Harris model with unadjusted support percentages
harris_evaluation <- evaluate_model(model_date_harris, data_test_harris)

# Evaluate Trump model with unadjusted support percentages
trump_evaluation <- evaluate_model(model_date_trump, data_test_trump)

# Augment data with model predictions
model_harris_adj <- readRDS(here::here("models/MLR_adjusted_harris.rds"))
model_trump_adj <- readRDS(here::here("models/MLR_adjusted_trump.rds"))

harris_2024_national_adj <- harris_2024_national_adj %>% mutate(
  fitted_date = predict(model_harris_adj)
)

trump_2024_national_adj <- trump_2024_national_adj %>% mutate(
  fitted_date = predict(model_trump_adj)
)

# Unadjusted datasets
model_harris <- readRDS(here::here("models/MLR_harris.rds"))
model_trump <- readRDS(here::here("models/MLR_trump.rds"))

harris_2024_national <- harris_2024_national %>% mutate(
  fitted_date = predict(model_harris)
)

trump_2024_national <- trump_2024_national %>% mutate(
  fitted_date = predict(model_trump)
)

# evaluate models without using training and testing data
harris_eval_adj <- evaluate_model_adj(model_harris_adj, harris_2024_national_adj)

harris_eval <- evaluate_model(model_harris, harris_2024_national)

trump_eval_adj <- evaluate_model_adj(model_trump_adj, trump_2024_national_adj)

trump_eval <- evaluate_model(model_trump, trump_2024_national)

#### Logistic regression for national support ####

combined_national <- full_join(
  trump_2024_national,
  harris_2024_national,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_national <- combined_national %>% mutate(
  harris_win = case_when(
    pct.y >= pct.x ~ 1,
    pct.y < pct.x ~ 0
  )
)

# repeat for adjusted support levels
combined_national_adj <- full_join(
  trump_2024_national_adj,
  harris_2024_national_adj,
  by = c(
    "pollster", "has_sponsor", "pollscore",
    "transparency_score", "sample_size", "end_date", "state", "cycle"
  )
)

combined_national_adj <- combined_national_adj %>% mutate(
  harris_win = case_when(
    pct_adj.y >= pct_adj.x ~ 1,
    pct_adj.y < pct_adj.x ~ 0
  )
)

# logistic model to predict the popular vote
model_logistic <- readRDS(here::here("models/Logistic_model.rds"))

model_logistic_adj <- readRDS(here::here("models/Logistic_model_adjusted.rds"))

# include predicted outcomes to the dataframes
combined_national <- combined_national %>%
  mutate(
    harris_win_prob = round(100 * predict(model_logistic, type = "response"), 2)
  )

combined_national_adj <- combined_national_adj %>%
  mutate(
    harris_win_prob = round(100 * predict(model_logistic_adj,
      type = "response"
    ), 2)
  )

#### Evaluate the Model ####

# ROC Curve and AUC
roc_obj <- roc(
  combined_national$harris_win,
  combined_national$harris_win_prob
)
auc_value <- auc(roc_obj)

roc_obj_adj <- roc(
  combined_national_adj$harris_win,
  combined_national_adj$harris_win_prob
)
auc_value_adj <- auc(roc_obj_adj)
```

For the analysis of the national popular vote, the logistic model of adjusted polls possess an AUC value of 0.98, and the logistic model of unadjusted polls have an AUC of 0.94, indicating high reliability of categorization. Their respective ROC plots are presented as @fig-log and @fig-logadj.

RMSE values and R squared values were calculated for the multiple linear regression models predicting popular vote. The RMSE values for all four models, adjusted and unadjusted support levels of the two candidates, are greater than 1, indicating that the fitted values and the support levels of actual polls are more than 1 percent apart on average, which is significant when analyzing popular vote. Moreover, all R squared values other than the model for the adjusted support rate of Trump, are smaller than 0.3, indicating the inability of the models to account for the variance of national support rates in the data. The model for the adjusted support rate of Trump has a R squared value of 0.69, indicating a relatively high proportion of variance in polled support rates being explained by the multiple linear regression model. This was expected, as the predictors such as a poll's end date and transparency does not necessarily correlate with the outcome of the polls. More predictive variables used by FiveThirtyEight [@5383] for their predictions are unavailable in the raw dataset, which is a key limitation of the data.

## Swing states

@tbl-result shows the average likelihood percentage of Harris winning each swing state, calculated by using the logistic model on 1,000 hypothetical data points for each state, generated with the normal distribution of predictors and setting the end_date to election day. It indicates that Harris has a greater probability to win in Pennsylvania, Georgia, and Nevada, while Trump is more likely to win in North Carolina, Michigan, Arizona, and Wisconsin. The electoral votes in each state is also presented.

#### Results ####
# To predict the outcome on election day, generate 1000 hypothetical data points
# using the election day as end_date and the normal curve of other input
# variables, mimicing all the situations that may happen on election day
hypothetical_data_Pennsylvania <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Pennsylvania$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Pennsylvania$transparency_score),
    sd = sd(combined_Pennsylvania$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Pennsylvania$sample_size),
    sd = sd(combined_Pennsylvania$sample_size)
  )
)

hypothetical_data_Georgia <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Georgia$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Georgia$transparency_score),
    sd = sd(combined_Georgia$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Georgia$sample_size),
    sd = sd(combined_Georgia$sample_size)
  )
)

hypothetical_data_North_Carolina <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_North_Carolina$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_North_Carolina$transparency_score),
    sd = sd(combined_North_Carolina$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_North_Carolina$sample_size),
    sd = sd(combined_North_Carolina$sample_size)
  )
)

hypothetical_data_Michigan <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Michigan$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Michigan$transparency_score),
    sd = sd(combined_Michigan$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Michigan$sample_size),
    sd = sd(combined_Michigan$sample_size)
  )
)

hypothetical_data_Arizona <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Arizona$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Arizona$transparency_score),
    sd = sd(combined_Arizona$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Arizona$sample_size),
    sd = sd(combined_Arizona$sample_size)
  )
)

hypothetical_data_Wisconsin <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Wisconsin$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Wisconsin$transparency_score),
    sd = sd(combined_Wisconsin$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Wisconsin$sample_size),
    sd = sd(combined_Wisconsin$sample_size)
  )
)

hypothetical_data_Nevada <- tibble(
  end_date = as.Date("2024-11-05"),
  has_sponsor = rbinom(1000, size = 1, prob = mean(combined_Nevada$has_sponsor)),
  transparency_score = rnorm(1000,
    mean = mean(combined_Nevada$transparency_score),
    sd = sd(combined_Nevada$transparency_score)
  ),
  sample_size = rnorm(1000,
    mean = mean(combined_Nevada$sample_size),
    sd = sd(combined_Nevada$sample_size)
  )
)

# run the corresponding logistic regression on the hypothetical data of each
# state and average the outcomes to find the expected win rate of Harris
# on election day
hypothetical_data_Pennsylvania <- hypothetical_data_Pennsylvania %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Pennsylvania, newdata = ., type = "response"), 2)
  )

hypothetical_data_Georgia <- hypothetical_data_Georgia %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Georgia, newdata = ., type = "response"), 2)
  )

hypothetical_data_North_Carolina <- hypothetical_data_North_Carolina %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_North_Carolina, newdata = ., type = "response"), 2)
  )

hypothetical_data_Michigan <- hypothetical_data_Michigan %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Michigan, newdata = ., type = "response"), 2)
  )

hypothetical_data_Arizona <- hypothetical_data_Arizona %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Arizona, newdata = ., type = "response"), 2)
  )

hypothetical_data_Wisconsin <- hypothetical_data_Wisconsin %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Wisconsin, newdata = ., type = "response"), 2)
  )

hypothetical_data_Nevada <- hypothetical_data_Nevada %>%
  mutate(
    harris_win_prob = round(100 * predict(logistic_Nevada, newdata = ., type = "response"), 2)
  )

# produce a table of mean winning probabilities
mean_probabilities <- tibble(
  State = c("Pennsylvania", "Georgia", "North Carolina", "Michigan", "Arizona", "Wisconsin", "Nevada"),
  Harris_Win_Probability = c(
    round(mean(hypothetical_data_Pennsylvania$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Georgia$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_North_Carolina$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Michigan$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Arizona$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Wisconsin$harris_win_prob, na.rm = TRUE), 2),
    round(mean(hypothetical_data_Nevada$harris_win_prob, na.rm = TRUE), 2)
  )
)

# include electoral votes to determine who is in the lead
state_evs <- tibble::tribble(
  ~State, ~electoral_votes,
  "Alaska", 3,
  "Arizona", 11,
  "Arkansas", 6,
  "California", 55,
  "Colorado", 9,
  "Connecticut", 7,
  "Florida", 29,
  "Georgia", 16,
  "Illinois", 20,
  "Indiana", 11,
  "Iowa", 6,
  "Kansas", 6,
  "Louisiana", 8,
  "Maine", 2,
  "Maine CD-1", 1,
  "Maine CD-2", 1,
  "Maryland", 10,
  "Massachusetts", 11,
  "Michigan", 16,
  "Minnesota", 10,
  "Mississippi", 6,
  "Missouri", 10,
  "Montana", 3,
  "Nebraska", 2,
  "Nebraska CD-2", 1,
  "Nevada", 6,
  "New Hampshire", 4,
  "New Mexico", 5,
  "New York", 29,
  "North Carolina", 15,
  "Ohio", 18,
  "Oklahoma", 7,
  "Oregon", 7,
  "Pennsylvania", 20,
  "Rhode Island", 4,
  "South Carolina", 9,
  "South Dakota", 3,
  "Texas", 38,
  "Utah", 6,
  "Vermont", 3,
  "Virginia", 13,
  "Washington", 12,
  "Wisconsin", 10
)

mean_probabilities <- mean_probabilities %>%
  left_join(state_evs, by = "State")

#### Overall Probability of Winning ####
# Harris
# Create data frame of states
states <- data.frame(
  state = c("PA", "GA", "NC", "MI", "WI", "NV", "AZ"),
  prob = c(0.5551, 0.6801, 0.1033, 0.5151, 0.3792, 0.7679, 0),
  votes = c(20, 16, 15, 16, 10, 6, 11)
)

# Function to calculate probability of a specific combination
calc_combo_prob <- function(combo, states) {
  prob <- 1
  for (i in 1:nrow(states)) {
    if (i %in% combo) {
      prob <- prob * states$prob[i]
    } else {
      prob <- prob * (1 - states$prob[i])
    }
  }
  return(prob)
}

# Generate all possible combinations
n_states <- nrow(states)
total_prob <- 0

# Check each possible combination
for (i in 1:(2^n_states - 1)) {
  # Convert number to binary to get combination
  combo <- which(intToBits(i)[1:n_states] == 1)

  # Calculate total electoral votes for this combination
  votes <- sum(states$votes[combo])

  # If this combination has enough votes, add its probability
  if (votes >= 44) {
    prob <- calc_combo_prob(combo, states)
    total_prob <- total_prob + prob
  }
}

>>>>

library(tidyverse)
library(caret)
library(glmnet)
library(nnet)
library(stringr)
library(gt)
library(knitr)
library(kableExtra)
library(MatchIt)
library(arrow)
```

@@

I used the statistical programming language R [@R] for the graphing, analysis and presentation of the project as a whole. Caret [@caret], glmnet [@glmnet; @glmnet2; @glmnet3], nnet [@nnet], tidyverse [@tidyverse], car [@car] MatchIt [@matchit] and arrow [@arrow] were used directly in the analysis of the data. Stringr [@stringr], dplyr [@dplyr], gt [@gt], knitr [@knitr], kableExtra [@kextra] and styler [@styler] were used in the presentation and/or styling of the data, graphs, and paper as a whole.

@@

@@

To show the effects of PSM, @tbl-match provides a summary of balance for the match between the Quartile 1 and the Quartile 2 subset of CFPS 2020. The match was executed using caliper = 0.1 and a ratio of 2, matching 1041 Quartile 2 instances out of 1324 total instances with all 541 Quartile 1 instances. This matching procedure was repeated 6 times to compare each quartile with every other quartile for all 6 CFPS iterations from 2010 to 2020.

```{r}
#| label: tbl-match
#| tbl-cap: Summary of Balance for Quartile 1 vs Quartile 2 Match, 2020 Data
#| echo: false
#| message: false
#| warning: false

# Data computed from model_data script
data <- data.frame(
  Variables = c("distance", "Gender", "Work exp", "City Hukou", "Marital Status", "CCP", "East"),
  `Means Q1` = c(0.3016, 0.5712, 14.8817, 0.1072, 0.7375, 0.0240, 0.2810),
  `Means Q2` = c(0.2854, 0.5650, 14.1767, 0.1971, 0.7515, 0.0219, 0.3112),
  `Std Mean Diff` = c(0.3099, 0.0125, 0.0866, -0.2907, -0.0318, 0.0139, -0.0672),
  `Means Q1 Matched` = c(0.3016, 0.5712, 14.8817, 0.1072, 0.7375, 0.0240, 0.2810),
  `Means Q2 Matched` = c(0.3009, 0.5675, 14.7902, 0.1100, 0.7366, 0.0222, 0.2810),
  `Std Mean Diff M` = c(0.0139, 0.0075, 0.0112, -0.0090, 0.0021, 0.0121, 0.0000)
)

kable(
  data,
  format = "latex"
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position", "scale_down"), 
    position = "center",
    full_width = FALSE
  )
```
\FloatBarrier

@@

It is acknowledged that the regression models may be inconsistent due to the inconsistencies in the variables and measurements of different CFPS iterations. The incorporation of calculated variables is an additional factor impacting the models' ability to calculate returns to education.

@@

Finally, the Delta coefficient $\alpha_1$ from all paired comparisons across every CFPS iteration were compiled together to identify cross-quartile education return disparities over time.

As a test of model robustness, another set of regressions were ran without PSM to verify the overall significance of educational returns and the consistent pattern observed between quartiles. This will be discussed further in @sec-results.

@@

# Load all the models
# Define the path to the models folder
models_path <- here::here("models")

# List all RDS files in the models folder
model_files <- list.files(
  path = models_path,
  pattern = "\\.rds$",    
  full.names = TRUE       
)

# Load each RDS file and store in a list
models <- lapply(model_files, readRDS)

# Assign names to the list based on the filenames (without .rds)
names(models) <- tools::file_path_sans_ext(basename(model_files))

```

## Cross-quartile disparities in Educational Returns

```{r}
#| label: tbl-results
#| tbl-cap: Cross-Quartile Percentage Disparities of Returns to Education Over Time
#| echo: false
#| message: false
#| warning: false

library(knitr)
library(kableExtra)

# Data
data <- data.frame(
  `CFPS Year` = c(2020, 2018, 2016, 2014, 2012, 2010, "Temporal mean"),
  `Q1vsQ2` = c(0.9482, 0.2441, 0.7665, -2.3488, 1.2055, 0.4988, 0.2191),
  `Q1vsQ3` = c(1.9092, 2.3463, 1.2655, -1.1069, 0.9527, 3.0412, 1.4013),
  `Q1vsQ4` = c(4.1670, 3.9833, 4.5967, 0.6399, 3.1572, 5.6921, 3.7060),
  `Q2vsQ3` = c(1.2350, 1.7767, 1.2912, 0.2706, -0.0212, 2.1727, 1.1208),
  `Q2vsQ4` = c(3.1601, 3.3524, 3.9390, 3.3242, 3.7233, 4.9265, 3.7376),
  `Q3vsQ4` = c(2.2804, 3.7010, 3.1417, 3.1945, 4.4302, 2.1920, 3.1566),
  `Cross Sectional Mean` = c(2.2833, 2.5673, 2.5001, 0.6623, 2.2413, 3.0872, ' ')
)

# Create the table
kable(
  data,
  format = "latex"
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position", "scale_down"), 
    position = "center",
    full_width = FALSE
  )

```
\FloatBarrier
Note: In 2012, educational returns for Q1, Q2, and Q3 were insignificant. In 2010, educational returns for Q1 and Q2 were insignificant.

As @tbl-results shows, in most CFPS iterations, individuals from higher family income quartiles demonstrated a significant advantage in terms of returns to education compared to their counterparts in other quartiles. The greatest differences were observed when comparing the highest income quartile (Q4) to other quartiles. Over all CFPS iterations, individuals in the top quartile benefitted from 3.71% more gains in income from an additional year of education compared to the lowest quartile, 3.74% more than the second quartile, and 3.16% more than the third quartile on average. Comparatively, the differences between the other quartiles were less pronounced.

Upon evaluating cross-sectional means, no discernible upward or downward trend was identified. Other than the outlier year 2014, individuals from higher quartiles experienced around 2 to 3 percent greater educational returns on average compared to individuals from relatively lower quartiles. Thus, a general inference can be made: while those in the higher family income quartiles consistently enjoy significant benefits in terms of returns to education, this advantage has not expanded over time.

## Robustness testing

```{r}
#| label: tbl-robust
#| tbl-cap: Quartiles’ Percentage Returns to Education Over Time Without PSM
#| echo: false
#| message: false
#| warning: false

library(knitr)
library(kableExtra)

data <- data.frame(
  `CFPS Year` = c(2020, 2018, 2016, 2014, 2012, 2010),
  `Q1 Returns` = c('4.9507***', '4.1730**', '2.5671*', '4.4464**', '-1.0031', '0.1803'),
  `Q2 Returns` = c('5.8130***', '4.0239***', '3.0501***', '2.5357**', '0.0590', '1.1116'),
  `Q3 Returns` = c('7.3284***', '6.1466***', '4.1665***', '2.8027***', '0.4928', '3.9577***'),
  `Q4 Returns` = c('10.3891***', '9.1724***', '6.9051***', '6.3410***', '4.4870***', '8.0710***')
)

kable(
  data,
  format = "latex"
) %>%
  kable_styling(
    latex_options = c("striped", "hold_position", "scale_down"), 
    position = "center",
    full_width = FALSE
  )

```
\FloatBarrier
Significant codes: * = p < 0.05, ** = p < 0.01,  *** = p < 0.001

When directly regressing without matching, the significance of educational returns coefficients remained the same; the coefficients for Q1, Q2 and Q3 of CFPS2012 and the coefficients for Q1 and Q2 of CFPS2010 are the only ones insignificant at the 5% level. It is uncertain why educational returns are insignificant for these family income quartiles, both in the matched data and the unmatched data. It might be caused by the legacy data collection methods or the changes in variable recording that occurred after the 2012 iteration. Despite that, the outcome of the robustness test demonstrates the model's consistency.
